Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-11-20 manuscripts/supplement.Rmd
minor typos fixed

# Further sensitivity analysis (Fig S2) #

@@

```{r figure_S2, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/figureS2.csv") 
```
```{r S2_eml, eval=FALSE}
# not descriptive enough, supplement text documents this better
par <- levels(sensitivity_dat$parameters)
names(par) <- par
col.defs <- names(sensitivity_dat)
unit.defs <- list(models, par, "number", "number",  "number", "number")
col.classes <- sapply(sensitivity_dat, class)
eml_write("components/data/figureS2.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S2", file="components/data/README_S2.xml")
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S2." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S2." )
```

@@

## GP MCMC (Fig S3-4) ##

```{r figure_S3, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
gp_assessment_plots$traces_plot
```

```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red; note the posterior distributions are significantly more peaked than the priors, showing that the data has been informative and is not driven by the priors.", fig.width=4, fig.height=4}
gp_assessment_plots$posteriors_plot
write.csv(tgp_dat,"components/data/figureS3-4.csv")
```

@@

## Ricker Model MCMC (Fig S5-6) ##

```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood. ", fig.width=4}
ggplot(ricker_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model. Note that the model estimates a carrying capacity $K$ very close to the true equilibrium where most of the observations were made, but is less certain about the growth rate.", fig.width=4, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/figureS5-6.csv")
```

@@

## Myers Model MCMC (Fig S7-8) ##

```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
ggplot(myers_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model. Note that with more free parameters, the posteriors reflect greater uncertainty.  In particular, the parameter $\\theta$ includes values both above 2, resulting in a tipping point, and below 2, where no tipping point exists in the model.  Though the dynamic program will integrate over the full distribution, including those values corresponding to tipping points, the weight of the model lies in the region without tipping points.", fig.width=4, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/figureS7-8.csv")
```

@@

## Allen Model MCMC (Fig S9-10) ##

```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
ggplot(allen_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model. The Allen model is the structurally correct model.  Despite potential identfiability issues in distinguishing between the stable and unstable points ($K$ and $\\theta$ respectively), the posterior estimates successfully reflect both the the upper stable point ($K$), as well as the significant probabilty of a tipping point ($\\theta$) somewhere between $K$ and extinction (0).", fig.width=4, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/figureS9-10.csv")
```

>>>>

# Further sensitivity analysis (Fig S2 - 3) #

We perform 2 sensitivity analyses. The first focuses on illustrating the robustness of the
approach to the two parameters that most influence stochastic transitions across the tipping 
point: the position of the Allee threshold and the scale of the noise (Fig S2).

Changing the intensity of the stochasticity or the distance between stable and unstable steady states
does not impact the performance of the GP relative to the optimal solution
obtained from the true model and true parameters.  The parametric models are
more sensitive to this difference.  Large values of $\sigma$ relative to the distance between the stable and unstable
point increases the chance of a stochastic transition below the tipping point.  More precisely,
if we let $L$ be the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as

$\exp\left(\frac{L^2}{\sigma^2}\right)$

(see @Gardiner2009 or @Mangel2006 for the derivation).

Thus, the impact of using a model that  underestimates the risk of harvesting
beyond the critical point is considerable, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable steady state relative to $\sigma$, the chance
of a transition becomes vanishingly small and all models can be estimated 
near-optimally. Models that underestimate the cost incurred by population
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficiently small.
Fig S2 shows the net present value of managing under teh GPDP remains
close to the optimal value (ratio of 1), despite varying across either
noise level or the the position of the allee threshold.

```{r figure_S2, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).  Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect.", cache=FALSE}
plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
  geom_point(size=2) + ylab("Net present value") + 
  xlab("Level of growth stochasticity")
plot_allees <- ggplot(vary_allee, aes(allee, value)) + 
  geom_point(size=2)+ ylab("Net present value") + 
  xlab("Allee threshold stock size")
multiplot(plot_sigmas, plot_allees, cols=2) 
write.csv(vary_sigma, "components/data/figureS2a.csv")
write.csv(vary_allee, "components/data/figureS2b.csv")
```

@@

```{r figure_S3, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/figureS3.csv") 
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S3." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S3." )
```

@@

## GP MCMC (Fig S4-5) ##

```{r figure_S4, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
gp_assessment_plots$traces_plot
```

```{r figure_S5, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red; note the posterior distributions are significantly more peaked than the priors, showing that the data has been informative and is not driven by the priors.", fig.width=4, fig.height=4}
gp_assessment_plots$posteriors_plot
write.csv(tgp_dat,"components/data/figureS4-5.csv")
```

@@

## Ricker Model MCMC (Fig S6-7) ##

```{r figure_S6, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood. ", fig.width=4}
ggplot(ricker_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S7, fig.cap="Posteriors from the MCMC estimate of the Ricker model. Note that the model estimates a carrying capacity $K$ very close to the true equilibrium where most of the observations were made, but is less certain about the growth rate.", fig.width=4, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/figureS6-7.csv")
```

@@

## Myers Model MCMC (Fig S8-9) ##

```{r figure_S8, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
ggplot(myers_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S9, fig.cap="Posterior distributions from the MCMC estimates of the Myers model. Note that with more free parameters, the posteriors reflect greater uncertainty.  In particular, the parameter $\\theta$ includes values both above 2, resulting in a tipping point, and below 2, where no tipping point exists in the model.  Though the dynamic program will integrate over the full distribution, including those values corresponding to tipping points, the weight of the model lies in the region without tipping points.", fig.width=4, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/figureS8-9.csv")
```

@@

## Allen Model MCMC (Fig S10-11) ##

```{r figure_S10, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps). stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
ggplot(allen_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S11, fig.cap="Posteriors from the MCMC estimate of the Allen model. The Allen model is the structurally correct model.  Despite potential identfiability issues in distinguishing between the stable and unstable points ($K$ and $\\theta$ respectively), the posterior estimates successfully reflect both the the upper stable point ($K$), as well as the significant probabilty of a tipping point ($\\theta$) somewhere between $K$ and extinction (0).", fig.width=4, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/figureS10-11.csv")
```

### 2014-11-20 manuscripts/supplement.tex
minor typos fixed

While the data can be regenerated using the code provided, for
convenience CSV files of the data shown in each graph are made available
on Dryad.

@@

\section{Further sensitivity analysis (Fig
S2)}\label{further-sensitivity-analysis-fig-s2}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S2-1.pdf}
\caption{Sensitivity Analysis. Histograms shows the ratio of the
realized net present value derived when managing under the GPDP over the
optimal value given the true model and true parameters. Values of 1
indicate optimal performance. Columns indicate different models, rows
different noise levels, and colors indicate the parameter set used.
Grouped over stochastic replicates applying the contol policy and
stochastic replicates of training data generated from the model
indicated, see raw data for details. Randomly chosen parameter values
for the models shown in tables below.}
\end{figure}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\centering
theta
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.A}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.103
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
7.949
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
2.288
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.B}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.485
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
9.775
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
3.524
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Allen models in Figure
S2.}
\end{longtable}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
C
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.C}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.769
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.46
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.301
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.D}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
2.075
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.95
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.915
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Myers models in Figure
S2.}
\end{longtable}

@@

\subsection{GP MCMC (Fig S3-4)}\label{gp-mcmc-fig-s3-4}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S3-1.pdf}
\caption{Traces from the MCMC estimates of the GP model show reasonable
mixing (no trend) and sampling rejection rate (no piecewise jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S4-1.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red; note the posterior distributions are
significantly more peaked than the priors, showing that the data has
been informative and is not driven by the priors.}
\end{figure}

@@

\subsection{Ricker Model MCMC (Fig
S5-6)}\label{ricker-model-mcmc-fig-s5-6}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S5-1.pdf}
\caption{Traces from the MCMC estimates of the Ricker model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S6-1.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model. Note
that the model estimates a carrying capacity $K$ very close to the true
equilibrium where most of the observations were made, but is less
certain about the growth rate.}
\end{figure}

@@

\subsection{Myers Model MCMC (Fig
S7-8)}\label{myers-model-mcmc-fig-s7-8}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S7-1.pdf}
\caption{Traces from the MCMC estimates of the Myers model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S8-1.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model. Note that with more free parameters, the posteriors reflect
greater uncertainty. In particular, the parameter $\theta$ includes
values both above 2, resulting in a tipping point, and below 2, where no
tipping point exists in the model. Though the dynamic program will
integrate over the full distribution, including those values
corresponding to tipping points, the weight of the model lies in the
region without tipping points.}
\end{figure}

@@

\subsection{Allen Model MCMC (Fig
S9-10)}\label{allen-model-mcmc-fig-s9-10}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S9-1.pdf}
\caption{Traces from the MCMC estimates of the Allen model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S10-1.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model. The Allen
model is the structurally correct model. Despite potential
identfiability issues in distinguishing between the stable and unstable
points ($K$ and $\theta$ respectively), the posterior estimates
successfully reflect both the the upper stable point ($K$), as well as
the significant probabilty of a tipping point ($\theta$) somewhere
between $K$ and extinction (0).}
\end{figure}

@@

>>>>

While the data can be regenerated using the code provided, for
convenience CSV files of the data shown in each graph are made available
on Dryad, along with the source \texttt{.Rmd} files for the manuscript
and supplement that document them.

@@

\section{Further sensitivity analysis (Fig S2 -
3)}\label{further-sensitivity-analysis-fig-s2---3}

We perform 2 sensitivity analyses. The first focuses on illustrating the
robustness of the approach to the two parameters that most influence
stochastic transitions across the tipping point: the position of the
Allee threshold and the scale of the noise (Fig S2).

Changing the intensity of the stochasticity or the distance between
stable and unstable steady states does not impact the performance of the
GP relative to the optimal solution obtained from the true model and
true parameters. The parametric models are more sensitive to this
difference. Large values of $\sigma$ relative to the distance between
the stable and unstable point increases the chance of a stochastic
transition below the tipping point. More precisely, if we let $L$ be the
distance between the stable and unstable steady states, then the
probability that fluctuations drive the population across the unstable
steady state scales as

$\exp\left(\frac{L^2}{\sigma^2}\right)$

(see Gardiner (2009) or Mangel (2006) for the derivation).

Thus, the impact of using a model that underestimates the risk of
harvesting beyond the critical point is considerable, since this such a
situation occurs more often. Conversely, with large enough distance
between the optimal escapement and unstable steady state relative to
$\sigma$, the chance of a transition becomes vanishingly small and all
models can be estimated near-optimally. Models that underestimate the
cost incurred by population sizes fluctuating significantly below the
optimal escapement level will not perform poorly as long as those
fluctuations are sufficiently small. Fig S2 shows the net present value
of managing under teh GPDP remains close to the optimal value (ratio of
1), despite varying across either noise level or the the position of the
allee threshold.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S2-1.pdf}
\caption{The effect of increasing noise or decreasing Allee threshold
levels on the net present value of the fishery when managed under the
GPDP, relative to managing under the true model (with known parameters).
Other than the focal parameter (stochasticity, Allee threshold), other
parameters are held fixed as above to illustrate this effect.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S3-1.pdf}
\caption{Sensitivity Analysis. Histograms shows the ratio of the
realized net present value derived when managing under the GPDP over the
optimal value given the true model and true parameters. Values of 1
indicate optimal performance. Columns indicate different models, rows
different noise levels, and colors indicate the parameter set used.
Grouped over stochastic replicates applying the contol policy and
stochastic replicates of training data generated from the model
indicated, see raw data for details. Randomly chosen parameter values
for the models shown in tables below.}
\end{figure}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\centering
theta
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.A}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.103
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
7.949
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
2.288
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.B}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.485
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
9.775
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
3.524
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Allen models in Figure
S3.}
\end{longtable}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
C
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.C}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.769
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.46
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.301
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.D}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
2.075
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.95
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.915
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Myers models in Figure
S3.}
\end{longtable}

@@

\subsection{GP MCMC (Fig S4-5)}\label{gp-mcmc-fig-s4-5}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S4-1.pdf}
\caption{Traces from the MCMC estimates of the GP model show reasonable
mixing (no trend) and sampling rejection rate (no piecewise jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S5-1.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red; note the posterior distributions are
significantly more peaked than the priors, showing that the data has
been informative and is not driven by the priors.}
\end{figure}

@@

\subsection{Ricker Model MCMC (Fig
S6-7)}\label{ricker-model-mcmc-fig-s6-7}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S6-1.pdf}
\caption{Traces from the MCMC estimates of the Ricker model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S7-1.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model. Note
that the model estimates a carrying capacity $K$ very close to the true
equilibrium where most of the observations were made, but is less
certain about the growth rate.}
\end{figure}

@@

\subsection{Myers Model MCMC (Fig
S8-9)}\label{myers-model-mcmc-fig-s8-9}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S8-1.pdf}
\caption{Traces from the MCMC estimates of the Myers model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S9-1.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model. Note that with more free parameters, the posteriors reflect
greater uncertainty. In particular, the parameter $\theta$ includes
values both above 2, resulting in a tipping point, and below 2, where no
tipping point exists in the model. Though the dynamic program will
integrate over the full distribution, including those values
corresponding to tipping points, the weight of the model lies in the
region without tipping points.}
\end{figure}

@@

\subsection{Allen Model MCMC (Fig
S10-11)}\label{allen-model-mcmc-fig-s10-11}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S10-1.pdf}
\caption{Traces from the MCMC estimates of the Allen model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S11-1.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model. The Allen
model is the structurally correct model. Despite potential
identfiability issues in distinguishing between the stable and unstable
points ($K$ and $\theta$ respectively), the posterior estimates
successfully reflect both the the upper stable point ($K$), as well as
the significant probabilty of a tipping point ($\theta$) somewhere
between $K$ and extinction (0).}
\end{figure}

@@

Gardiner, C. 2009. Stochastic Methods: A Handbook for the Natural and
Social Sciences (Springer Series in Synergetics). Page 447. Springer.

Mangel, M. 2006. The Theoretical Biologist's Toolbox: Quantitative
Methods for Ecology and Evolutionary Biology. Cambridge University
Press.
